---
title: "Movement-tracking of cognitive processes: A tutorial using mousetrap"
author: "Dirk Wulff, Pascal Kieslich, Felix Henninger, Jonas Haslbeck & Michael Schulte-Mecklenbeck"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 1
---

# Introduction

This tutorial illustrates how to use the mousetrap package to process, analyze and visualize movement trajectories. It accompanies the paper of the same name.

The R code in this tutorial performs all analyses and creates all trajectory data related figures that are presented in the paper.

After running the initialization section, all subsequent sections are self-contained. That is, each section can be run independently, as it prepares the data needed for its analysis and plot.

# Initialization

```{r setup, include = FALSE}
# Set general chunk options
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  # Set figure aspect ratio based on defaults from figure export function
  fig.asp = 0.64
  )
```


## Load libraries required in this tutorial
```{r}
library(mousetrap)
library(tidyverse)
library(patchwork)
library(psych)
library(MBESS)
library(viridis)
library(afex)
library(osfr)
```

## Set custom ggplot2 theme and colors
```{r}
theme_set(theme_minimal())
custom_colors <- cividis(5)
custom_colors_3 <- custom_colors[c(1, 4, 5)]
```

```{r, include = FALSE}
# Create custom figure export functions and set defaults

# Specify if figures created in this tutorial should also be saved as files
save_figures <- FALSE

# Specify folder in current working directory where figures should be saved
figure_path <- "3_figures"

# Create export function for ggsave
export_ggsave <- function(
  filename, 
  path = figure_path,
  save = save_figures,
  width = 6.3, height = 4.02, unit = "in", dpi = 300
){
  
  # Only export figures if specified in general settings
  if (save) {
    ggsave(
      filename = file.path(path, filename),
      width = width, height = height, unit = unit, dpi = dpi
    )
  }
}

# Create export function for standard devices
export_device <- function(
  filename, 
  path = figure_path,
  save = save_figures,
  current_plot,
  device_function,
  ...
){
  device_function(file.path(figure_path, filename), ...)
  replayPlot(current_plot)
  dev.off()
}
```


# Figure 1: Setup

## Preprocess data
```{r}
# Import and preprocess mouse-tracking data
mt_data <- KH2017_raw %>%
  filter(correct == 1) %>%
  mt_import_mousetrap() %>%
  mt_remap_symmetric() %>%
  mt_align_start(start = NULL) %>%
  mt_subset(mt_id %in% c("id0013", "id0030", "id0033"))

# Specify coordinates of buttons that should be plotted
rectangles <- matrix(
  c(
    -840, 525, 350, -170,
    840, 525, -350, -170
  ),
  ncol = 4, byrow = TRUE
)

# Specify button labels
button_labels <- data.frame(
  label = c("Whale", "Mammal", "Fish"),
  xpos = c(0, -665, 665),
  ypos = c(-440 + 85 + 40, 440, 440)
)
```

## Create plot
```{r}
mt_plot(mt_data, return_type = "mapping") +
  mt_plot_add_rect(rectangles) +
  coord_cartesian(xlim = c(-840, 840), ylim = c(-525, 525), expand = FALSE) +
  geom_path(aes(color = mt_id), size = 2) +
  geom_text(
    aes(x = xpos, y = ypos, label = label), data = button_labels, size = 3.2
    ) +
  theme(legend.position = "none") +
  scale_color_manual(values = custom_colors_3) +
  labs(x = NULL, y = NULL)
```

```{r, include = FALSE}
export_ggsave("figure_1_setup.pdf")
```


## Descriptives for raw trajectories
```{r}
mt_data <- KH2017_raw %>%
  mt_import_mousetrap()
summary(mt_count(mt_data$trajectories))
summary(mt_data$data$response_time)
```


# Figure 4: Resampling

## Preprocess data
```{r}
# Import and preprocess mouse-tracking data
mt_data <- KH2017_raw %>%
  filter(correct == 1) %>%
  mt_import_mousetrap() %>%
  mt_remap_symmetric(remap_xpos = "no") %>%
  mt_subset(mt_id %in% c("id0013", "id0030", "id0033"))

# Only select every second trajectory position
mt_data$trajectories <- mt_data$trajectories[, seq(1, 151, 2), ]
```

## Create plot
```{r}
# Plot raw trajectories
p1 <- mt_plot(mt_data, return_type = "mapping") +
  geom_path(color = "white", alpha = 1, size = .3, show.legend = FALSE) +
  geom_path(aes(color = mt_id), alpha = .7, size = .3, show.legend = FALSE) +
  geom_point(color = "white", alpha = 1, show.legend = FALSE) +
  geom_point(aes(color = mt_id), alpha = .7, show.legend = FALSE) +
  coord_cartesian(xlim = c(-840, 840), ylim = c(-525, 525), expand = FALSE) +
  scale_color_manual(values = custom_colors_3) +
  labs(x = NULL, y = NULL) + 
  theme(axis.text.x = element_blank())

# Remap trajectories, align their start position and plot them
mt_data <- mt_remap_symmetric(mt_data)
mt_data <- mt_align_start(mt_data, start = NULL)
p2 <- mt_plot(mt_data, return_type = "mapping") +
  geom_path(color = "white", alpha = 1, size = .3, show.legend = FALSE) +
  geom_path(aes(color = mt_id), alpha = .7, size = .3, show.legend = FALSE) +
  geom_point(color = "white", alpha = 1, show.legend = FALSE) +
  geom_point(aes(color = mt_id), alpha = .7, show.legend = FALSE) +
  coord_cartesian(xlim = c(-840, 840), ylim = c(-525, 525), expand = FALSE) +
  scale_color_manual(values = custom_colors_3) +
  labs(x = NULL, y = NULL) + 
  theme(axis.text.x = element_blank(), axis.text.y = element_blank())

# Time normalize and plot trajectories
mt_data <- mt_time_normalize(mt_data)
p3 <- mt_plot(mt_data, use = "tn_trajectories", return_type = "mapping") +
  geom_path(color = "white", alpha = 1, size = .3, show.legend = FALSE) +
  geom_path(aes(color = mt_id), alpha = .7, size = .3, show.legend = FALSE) +
  geom_point(color = "white", alpha = 1, show.legend = FALSE) +
  geom_point(aes(color = mt_id), alpha = .7, show.legend = FALSE) +
  coord_cartesian(xlim = c(-840, 840), ylim = c(-525, 525), expand = FALSE) +
  scale_color_manual(values = custom_colors_3) +
  labs(x = NULL, y = NULL)

# Length normalize and plot trajectories
mt_data <- mt_length_normalize(mt_data)
p4 <- mt_plot(mt_data, use = "ln_trajectories", return_type = "mapping") +
  geom_path(color = "white", alpha = 1, size = .3, show.legend = FALSE) +
  geom_path(aes(color = mt_id), alpha = .7, size = .3, show.legend = FALSE) +
  geom_point(color = "white", alpha = 1, show.legend = FALSE) +
  geom_point(aes(color = mt_id), alpha = .7, show.legend = FALSE) +
  coord_cartesian(xlim = c(-840, 840), ylim = c(-525, 525), expand = FALSE) +
  scale_color_manual(values = custom_colors_3) +
  labs(x = NULL, y = NULL) + 
  theme(axis.text.y = element_blank())

(p1 + p2) / (p3 + p4) & plot_annotation(tag_levels = "A")
```

```{r, include = FALSE}
export_ggsave("figure_4_resampling.pdf")
```


# Figure 5: Outliers

## Preprocess data
```{r}
# Preprocess mouse-tracking data
mt_data <- KH2017 %>%
  mt_time_normalize() %>%
  mt_length_normalize() %>%
  mt_map() %>%
  mt_standardize(use = "prototyping", use_variables = "min_dist")

# Classify outliers
mt_data$data$outlier <- ifelse(
  mt_data$prototyping$z_min_dist > 2,
  "Distance > 2 SD", "Distance <= 2 SD"
)
```

## Create plot
```{r}
mt_plot(mt_data, color = "outlier", return_type = "mapping") +
  geom_path(aes(alpha = outlier)) +
  theme(legend.position = c(.14, .2), legend.background = element_blank()) +
  scale_color_manual(name = "", values = custom_colors[c(3, 1)]) +
  scale_alpha_manual(name = "", values = c(.08, .6)) +
  labs(x = NULL, y = NULL) +
  coord_cartesian(xlim = c(-950, 950), ylim = c(-150, 1050), expand = FALSE)
```

```{r, include = FALSE}
export_ggsave("figure_5_outliers.pdf")
export_ggsave("figure_5_outliers.png")
```


# Figure 6: Trajectory indices

## Preprocess data
```{r}
# Import and preprocess mouse-tracking data and compute indices
mt_data <- KH2017_raw %>%
  filter(correct == 1) %>%
  mt_import_mousetrap() %>%
  mt_remap_symmetric() %>%
  mt_align_start() %>%
  mt_time_normalize() %>%
  mt_derivatives() %>%
  mt_measures() %>%
  mt_sample_entropy()

# Calculate movement time
mt_data$measures$movement <- mt_data$measures$RT - mt_data$measures$idle_time

# Calculate motor pauses
mt_data$measures$motor_pauses <-
  mt_data$measures$idle_time - mt_data$measures$initiation_time

# Select measures and specify labels
measures <- mt_data$measures[, c(
  "MAD", "MD_above", "AD", "AUC", "xpos_flips",
  "xpos_reversals", "sample_entropy", "RT", "initiation_time",
  "motor_pauses", "movement"
)]

labels <- c(
  "MAD", expression(MD[above]), "AD", "AUC", "Flips",
  "Reversals", "Sample entropy", "RT", "Initiation time",
  "Motor pauses", "Movement"
)

# Compute Pearson and Spearman rank correlations
cors <- measures %>% cor()
cors_rank <- measures %>% cor(method = "spearman")

# Compute Cohen's d with confidence interval per measure
cond <- mt_data$data$Condition
effects <- sapply(measures, function(x) cohen.d(x, cond == "Atypical")$cohen.d)

# Setup colored correlations grid
cors_combined <- cors
cors_combined[upper.tri(cors_combined)] <- cors_rank[upper.tri(cors_combined)]
cols <- rev(cividis(201))
offset <- c(rep(0, 4), rep(.5, 3), rep(1, 4))
pos <- expand.grid(x = (1:11) + offset, y = (1:11) + offset) %>%
  mutate(
    cor = as.vector(t(cors_combined)),
    col = (cols)[round(c(cor) * 100) + 101]
  ) %>%
  filter(x != y) %>%
  mutate(
    y = 1 + max(y) - y
  )
```

## Create plot
```{r, fig.asp = 0.695}
# Setup two pane plot layout
layout(matrix(1:2, ncol = 2), width = c(.6, .4))
par(mar = c(1, 5, 5, 1))

# Create correlations plot
w <- .5
plot.new()
plot.window(xlim = range(pos$x) + c(-w, w), ylim = range(pos$y) + c(-w, w))
rect(
  pos$x - w, pos$y - w, pos$x + w, pos$y + w,
  col = pos$col, border = NA
)
text(pos$x, pos$y, labels = round(pos$cor, 2),
     col = ifelse(pos$cor != 1, "white", cols[201]), cex = .5, font = 1)
mtext(labels, side = 2, at = unique(pos$y) %>% sort(decreasing = T),
      las = 1, adj = 1, cex = .75)
mtext(rev(labels), side = 3, at = unique(pos$y) %>% sort(decreasing = T),
      las = 2, adj = 0, cex = .75)
mtext("A", side = 3, at = -3, cex = 1.15, line = 2.5)

# Create Cohen's d plot
ypos <- unique(pos$y) %>% sort(decreasing = TRUE)
cols <- rev(cividis(100))
plot.new()
plot.window(xlim = c(-.2, .8), ylim = range(ypos) + c(-.5, .5))
d_lines <- sapply(seq(-.2, .8, .1), function(x)
  lines(c(x, x), range(ypos) + c(-.5, .5), lty = 2, lwd = .5)
  )
lines(c(0, 0), range(ypos) + c(-.5, .5), lwd = 2)
for (i in 1:length(ypos)) lines(effects[c(1, 3), i], ypos[c(i, i)], lwd = 2)
points(effects[2, ], ypos, pch = 15, cex = 1.5,
       col = cols[(effects[2, ] + .3) * 100 %>% round()])
labs <- sapply(seq(-.2, .8, .1) %>% round(1), function(x)
  str_sub(x, nchar(x) - 1, nchar(x))
  )
mtext(labs, at = seq(-.2, .8, .1), side = 3, cex = .8)
mtext(expression(paste("Cohen's ", italic(d))), side = 3, line = 2)
mtext(labels,
  side = 2, at = unique(pos$y) %>% sort(decreasing = T),
  las = 1, adj = 1, cex = .75, line = .5
)
mtext("B", side = 3, at = -.8, cex = 1.15, line = 2.5)

# Store plot for potential export
current_plot <- recordPlot()
```

```{r, include = FALSE}
export_device(
  "figure_6_indices.pdf",
  current_plot = current_plot,
  device_function = pdf,
  height = 4.02 * 1.1, width = 7 * 1.1
  )
```


## Average correlations between different types of measures
```{r}
# Curvature and complexity indices
cors[1:4, 5:7] %>% mean()

# Curvature and temporal indices
cors[1:4, c(8, 10, 11)] %>% mean()

# Complexity and temporal indices
cors[5:7, c(8, 10, 11)] %>% mean()
```

## Principal components analysis
```{r}
# One factor
pca(measures[, -8], nfactors = 1)[["communality"]] %>% mean()

# Five factors
pca(measures[, -8], nfactors = 5)[["communality"]] %>% mean()
```


# Figure 7: Homogeneity

## Preprocess data
```{r}
mt_data <- KH2017 %>%
  mt_time_normalize() %>%
  mt_length_normalize()
```

## Create plot
```{r, dev = "png", fig.width = 6.3, fig.height = 8.04, fig.asp = 1.28, dpi = 200}
# Setup two pane plot layout
par(mfrow = c(2, 1))

# Heatmap of raw trajectories
mt_heatmap(
  mt_data,
  variable = mt_data$data$Condition == "Atypical",
  smooth_radius = .5,
  mean_image = .15,
  mean_color = .15,
  colors = c("white", cividis(7)[c(6, 1)]),
  xres = 2000, bounds = c(-960, -100, 960, 1080),
  verbose = FALSE
)
mtext("A", cex = 1.5, side = 3, at = 50, line = -1.5)

# Heatmap of smoothed differences
mt_diffmap(
  mt_data,
  condition = mt_data$data$Condition == "Typical",
  colors = c(cividis(7)[6], "white", cividis(7)[1]),
  xres = 1000, bounds = c(-960, -100, 960, 1080),
  smooth_radius = 20, n_shades = 10,
  verbose = FALSE
)
mtext("B", cex = 1.5, side = 3, at = 50, line = -1.5)

# Store plot for potential export
current_plot <- recordPlot()
```

```{r, include = FALSE}
export_device(
  "figure_7_homogeneity.png",
  current_plot = current_plot,
  device_function = png,
  width = 6.3, height = 8.04, unit = "in", res = 600
  )
```


# Figure 8: Clustering

## Preprocess data
```{r}
# Preprocess trajectory data
mt_data <- KH2017 %>%
  mt_length_normalize() %>%
  mt_cluster(use = "ln_trajectories") %>%
  mt_map(use = "ln_trajectories")

# Set colors
col <- custom_colors[1]
col2 <- "white"
```

## Create plot
```{r, dev = "png", fig.width = 42, fig.height = 42, fig.asp = 1}
# Setup three pane plot layout
layout(matrix(1:18, ncol = 3, byrow = TRUE), height = c(.2, rep(1, 5)))

# Prepare prototypes to display
prototypes <- mt_length_normalize(mt_prototypes, 20)
prototypes[, , 1] <-
  prototypes[, , 1] * -mean(mt_data$ln_trajectories[, 20, "xpos"])
prototypes[, , 2] <-
  prototypes[, , 2] * (mean(mt_data$ln_trajectories[, 20, "ypos"]) / 1.5)

# Compute percentages to display
tab1 <- table(mt_data$clustering$cluster)
tab1 <- round(tab1 / sum(tab1), 2) * 100
tab2 <- table(mt_data$prototyping$prototype)
tab2 <- round(tab2 / sum(tab2), 2) * 100

# Specify labels
txts <- c("Clustering", "Prototypes", "Prototype clustering")
txts2 <- c("A", "B", "C")
txts3 <- dimnames(mt_prototypes)[[1]]

# Plot column labels
# Note: setup of a figure of this size only works when opening graphics device
#        of sufficient size (see commented out png call above)
for (i in 1:3) {
  plot.new()
  par(mar = c(0, 0, 0, 0))
  plot.window(xlim = c(0, 1), c(0, .2))
  text(.02, .1, txts2[i], font = 1, cex = 12)
}

# Plot trajectories per row
for (i in 1:5) {

  # First column in row: Trajectories in cluster ----
  
  # Extract clustered trajectories
  current_traj <- mt_subset(mt_data, cluster == i, check = "clustering")
  
  # Plot individual trajectories
  mt_heatmap(
    current_traj,
    smooth_radius = 1,
    colors = c("white", col),
    bounds = c(-1000, -100, 1000, 1100),
    xres = 1000,
    verbose = FALSE
  )

  # Add aggregate trajectory
  x <- colMeans(current_traj$ln_trajectories[, , "xpos"]) / 2 + 500
  y <- (colMeans(current_traj$ln_trajectories[, , "ypos"]) + 100) / 2.05
  lines(x, y, lwd = 55, col = col2)
  points(x[c(1, length(x))], y[c(1, length(y))],
         bg = col, col = col2, cex = 17, pch = 21, lwd = 10)
  lines(x, y, lwd = 35, col = col)
  
  # Add percentage labels
  text(200, 120, paste0(tab1[i], "%"), cex = 10)
  
  
  # Second column in row: Prototype ----
  plot.new()
  plot.window(xlim = c(-1000, 1000), ylim = c(-100, 1100))
  points(prototypes[i, c(1, 20), 1], prototypes[i, c(1, 20), 2],
         bg = col, col = col2, cex = 17, pch = 21, lwd = 7)
  lines(prototypes[i, , 1], prototypes[i, , 2], lwd = 35, col = col)
  x <- ifelse(i <= 3, ifelse(i == 1, 50, ifelse(i == 2, 300, -150)), 0)
  text(x, 500, labels = txts3[i], col = "black", cex = 10)

  
  # Third column in row: Trajectories mapped on prototype ----
  
  # Extract trajectories mapped on prototype
  current_traj <- mt_subset(mt_data, prototype == i, check = "prototyping")
  
  # Plot individual trajectories
  mt_heatmap(
    current_traj,
    smooth_radius = 1,
    colors = c("white", col),
    bounds = c(-1000, -100, 1000, 1100),
    xres = 1000,
    verbose = FALSE
  )

  # Add aggregate trajectory
  x <- colMeans(current_traj$ln_trajectories[, , "xpos"]) / 2 + 500
  y <- (colMeans(current_traj$ln_trajectories[, , "ypos"]) + 100) / 2.05
  lines(x, y, lwd = 55, col = col2)
  points(x[c(1, length(x))], y[c(1, length(y))],
         bg = col, col = col2, cex = 17, pch = 21, lwd = 10)
  lines(x, y, lwd = 35, col = col)
  
  # Add percentage labels
  text(200, 120, paste0(tab2[i], "%"), cex = 10)
}

# Store plot for potential export
current_plot <- recordPlot()
```

```{r, include = FALSE}
export_device(
  "figure_8_clustering.png",
  current_plot = current_plot,
  device_function = png,
  width = 3000, height = 602 * 5.133, unit = "px"
  )
```

## Prototype frequency comparison
```{r}
# Prototype frequencies per condition
prototype_frequencies <- 
  table(
    mt_data$data$Condition,
    mt_data$prototyping$prototype_label
    )
prototype_frequencies

# Percentages
prototype_frequencies/
  c(table(mt_data$data$Condition))

# Chi-squared test of prototype frequency
prototype_chisq <-
  chisq.test(prototype_frequencies)
prototype_chisq

# Extract residuals
prototype_chisq$residuals
```


# Figure 9: Position & angle

## Preprocess data
```{r}
# Preprocess trajectory data
mt_data <- KH2017 %>%
  mt_time_normalize() %>%
  mt_angles(use = "tn_trajectories")

# Transform angle and position into long format
pos_angle_long <- mt_export_long(
  mt_data,
  use = "tn_trajectories",
  use_variables = c("steps", "xpos", "angle_v"),
  use2_variables = c("Condition", "subject_nr")
)

# Setup function that runs a mixed model per time step
mixed_model_per_step <- function(step, dv){
  current_data <-
    pos_angle_long %>%
    filter(steps == step) %>%
    filter(is.na(.data[[dv]])==FALSE)
  
  # Do not run model if there is no non-NA data
  if(nrow(current_data) == 0){
    current_model <- "not_run"
    current_p <- 1
    current_n <- 0
    
  # If there is data, count number of observations
  } else{
    current_desc <-
      current_data %>%
      group_by(Condition) %>%
      summarize(
        n = n(),
        sd = sd(.data[[dv]]),
        n_subjects = length(unique(subject_nr))
      )
    current_n <- sum(current_desc$n)
    
    # Only run model if there is data for at least two participants and 
    # if SD of variable is > 0 in each condition
    if((min(current_desc$n_subjects) > 2) & (min(current_desc$sd > 0))){
      current_model <- mixed(
        as.formula(paste(dv, "(1|subject_nr)+Condition", sep = "~")),
        data = current_data,
        progress = FALSE
      )
      current_p <- current_model$anova_table$`Pr(>F)`
      current_model <- "run"
    
    # Otherwise set p to 1
    } else{
      current_model <- "not_run"
      current_p <- 1
    }
    
  }
  
  return(
    tibble(
      steps = step,
      p = current_p,
      sig = p < .05,
      n = current_n,
      model = current_model
    )
  )
  
}

# Run mixed models for xpos and angle_v
mixed_models_xpos <-
  unique(pos_angle_long$steps) %>%
  map_dfr(mixed_model_per_step, dv = "xpos")

mixed_models_angle <-
  unique(pos_angle_long$steps) %>%
  map_dfr(mixed_model_per_step, dv = "angle_v")

# Retrieve significant differences
mixed_models_xpos$steps[mixed_models_xpos$sig]
mixed_models_angle$steps[mixed_models_angle$sig]
```

## Create plot
```{r, fig.asp = 0.5}
# Create plot for x positions
p1 <- mt_plot(
  mt_data,
  use = "tn_trajectories",
  x = "steps", y = "xpos",
  color = "Condition", alpha = .1
) +
  mt_plot_aggregate(
    mt_data,
    use = "tn_trajectories",
    x = "steps", y = "xpos",
    color = "Condition",
    size = 2,
    return_type = "geom"
  ) +
  scale_color_manual(values = custom_colors_3[c(1, 3)]) +
  labs(x = "Time step", y = "Position on horizontal axis (x)") +
  geom_text(
    label = "*", color = "black", size = 2,
    mapping = aes(x = steps, y = 1000),
    data = mixed_models_xpos %>% filter(sig)
  )+ 
  theme(legend.position = "none")


# Create plot for angles
p2 <- mt_plot(
  mt_data,
  use = "tn_trajectories",
  x = "steps", y = "angle_v",
  color = "Condition", alpha = .05
) +
  mt_plot_aggregate(
    mt_data,
    use = "tn_trajectories",
    x = "steps", y = "angle_v",
    color = "Condition",
    size = 2,
    return_type = "geom",
    .funs = ~ mean(.x, na.rm = TRUE)
  ) +
  scale_color_manual(values = custom_colors_3[c(1, 3)]) +
  scale_y_continuous(
    breaks = seq(-pi, pi, pi / 2),
    labels = scales::math_format(.x * pi, format = function(x) x / pi)
    ) +
  labs(x = "Time step", y = "Angle relative to vertical axis (y)") +
  geom_text(
    label = "*", color = "black", size = 2,
    mapping = aes(x = steps, y = pi),
    data = mixed_models_angle %>% filter(sig)
  )+ 
  theme(legend.position = "top")

p1 + p2 + plot_annotation(tag_levels = "A")
```

```{r, include = FALSE}
export_ggsave("figure_9_temporal.pdf", width = 7, height = 3.5)
export_ggsave("figure_9_temporal.png", width = 7, height = 3.5)
```


# Figure 10: Temporal averaging

## Preprocess data
```{r}
# Preprocess trajectory data
mt_data <- KH2017 %>%
  mt_length_normalize() %>%
  mt_map(use = "ln_trajectories") %>%
  mt_measures() %>%
  # Resample trajectories to allow averaging non-normalized trajectories
  mt_resample(step_size = 10, exact_last_timestamp = FALSE)

# Add data from other elements to data element
mt_data$data$RT <- mt_data$measures$RT
mt_data$data$prototype_label <- mt_data$prototyping$prototype_label

```

## Descriptives
```{r}
mt_aggregate(
  mt_data,
  use = "measures", use_variables = "RT",
  use2_variables = "Condition",
  .funs = "median"
  )

rt_freq <- table(cut(mt_data$data$RT, breaks = c(0, 1500.5, 2500.5, 5000.5, Inf)))
rt_freq
rt_freq / nrow(mt_data$data)
```

## Create plot
```{r, fig.asp = 0.86}
p1 <- mt_plot(
  mt_data,
  use = "rs_trajectories", x = "timestamps", y = "xpos",
  color = "Condition", alpha = .1,
  subset = RT <= 1500
  ) +
  scale_color_manual(values = cividis(3)[c(1, 3)]) +
  labs(x = "Time in ms", y = "Position (x)", subtitle = "RT < 1500") +
  theme_minimal() + 
  theme(legend.position = "none",
        plot.subtitle = element_text(size = 8, hjust = 0.5)) +
  coord_cartesian(ylim = c(-800, 800)) +
  mt_plot_aggregate(
    mt_data,
    use = "rs_trajectories", x = "timestamps", y = "xpos",
    color = "Condition",
    size = 2,
    return_type = "geom",
    subset = RT <= 1500
  ) 

p2 <- mt_plot(
  mt_data,
  use = "rs_trajectories", x = "timestamps", y = "xpos",
  color = "Condition", alpha = .1,
  subset = RT > 1500 & RT <= 2500
) +
  scale_color_manual(values = cividis(3)[c(1, 3)]) +
  labs(x = "Time in ms", y = "Position (x)", subtitle = "1500 < RT <= 2500") +
  theme_minimal() + 
  theme(legend.position = "none",
        plot.subtitle = element_text(size = 8, hjust = 0.5)) +
  coord_cartesian(ylim = c(-800, 800)) +
  mt_plot_aggregate(
    mt_data,
    use = "rs_trajectories", x = "timestamps", y = "xpos",
    color = "Condition",
    size = 2,
    return_type = "geom",
    subset = RT > 1500 & RT <= 2500
  ) +
  theme(axis.text.y = element_blank(), axis.title.y = element_blank())

p3 <- mt_plot(
  mt_data,
  use = "rs_trajectories", x = "timestamps", y = "xpos",
  color = "Condition", alpha = .1,
  subset = RT > 2500 & RT <= 5000
) +
  scale_color_manual(values = cividis(3)[c(1, 3)]) +
  labs(x = "Time in ms", y = "Position (x)", subtitle = "2500 < RT <= 5000") +
  theme_minimal() + 
  theme(legend.position = "none",
        plot.subtitle = element_text(size = 8, hjust = 0.5)) +
  coord_cartesian(ylim = c(-800, 800)) +
  mt_plot_aggregate(
    mt_data,
    use = "rs_trajectories", x = "timestamps", y = "xpos",
    color = "Condition",
    size = 2,
    return_type = "geom",
    subset = RT > 2500 & RT <= 5000
  )+
  theme(axis.text.y = element_blank(), axis.title.y = element_blank())


h <- ggplot(mt_data$data, aes(x = RT, fill = Condition, color = Condition)) +
  geom_density(alpha = .55) +
  guides(fill = guide_legend(override.aes = list(alpha=1))) +
  theme_minimal() +
  scale_fill_manual(values = cividis(3)[c(1, 3)]) +
  scale_color_manual(values = cividis(3)[c(1, 3)]) +
  xlim(0, 5000) +
  theme(legend.position = "top") +
  labs(x = "Response time in ms", y = "Density") +
  theme(axis.text.y = element_blank()) +
  theme(legend.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.key.size = unit(.8, "lines")) +
  geom_vline(xintercept = c(1500, 2500), linetype = "dashed", size = .5)

h2 <- ggplot(
  mt_data$data, aes(x = RT, fill = prototype_label, color = prototype_label)
  ) +
  geom_density(alpha = .55) +
  guides(fill = guide_legend(override.aes = list(alpha=1))) +
  theme_minimal() +
  scale_fill_manual(values = cividis(5), name = "Trajectory type") +
  scale_color_manual(values = cividis(5), name = "Trajectory type") +
  xlim(c(0, 5000)) +
  theme(legend.position = "top") +
  labs(x = "Response time in ms", y = "Density") +
  theme(axis.text.y = element_blank()) +
  theme(legend.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.key.size = unit(.8, "lines")) +
  geom_vline(xintercept = c(1500, 2500), linetype = "dashed", size = .5)

h / (p1 + p2 + p3) / h2 + plot_annotation(tag_levels = "A")
```

```{r, include = FALSE}
export_ggsave("figure_10_temporal_agg.pdf", width = 7, height = 6)
export_ggsave("figure_10_temporal_agg.png", width = 7, height = 6)
```


# Figure 11: Velocity & acceleration

## Preprocess data
```{r}
## Create prototype mapping using standard trajectory preprocessing first
mt_data <- KH2017 %>%
  mt_length_normalize() %>%
  mt_map(use = "ln_trajectories")
mt_data$data$prototype_label <- mt_data$prototyping$prototype_label

## Exclude potential phase without movement at beginning and end of trial
mt_data <- mt_data %>%
  mt_exclude_initiation(reset_timestamps = TRUE) %>%
  mt_exclude_finish() 

## Calculate derivatives and then time-normalize trajectories with
## dimensions set to all meaning that derivatives are also time-normalized
mt_data <- mt_data %>%
  mt_derivatives() %>%
  mt_time_normalize(dimensions = "all")
```

## Smoothing function
A similar function will be included in mousetrap package.

```{r}
smooth <- function(x, pos, sd = 2) {
  sm <- numeric(length(x))
  for (i in 1:length(x)) {
    w <- dnorm(pos, pos[i], sd = sd)
    sm[i] <- sum(x * w, na.rm = T) / sum(w, na.rm = T)
  }
  sm
}
```

## Create plot
```{r, fig.asp = 0.4375}
agg_traj <- mt_aggregate(
  mt_data,
  use = "tn_trajectories",
  use2_variables = "prototype_label", trajectories_long = TRUE
) %>%
  group_by(prototype_label) %>%
  mutate(
    sm_vel = smooth(vel, steps, sd = 3),
    sm_acc = smooth(acc, steps, sd = 6)
  ) %>%
  ungroup()

vel_plot <- ggplot(
  agg_traj,  aes(steps, sm_vel, col = prototype_label, fill = prototype_label)
  ) +
  facet_wrap(~prototype_label, ncol = 5) +
  geom_path(size = 2, show.legend = FALSE) +
  scale_color_manual(values = cividis(5)) +
  scale_fill_manual(values = cividis(5)) +
  labs(
    x = "Time step",
    y = "Average velocity"
  )

acc_plot <- ggplot(
  agg_traj,  aes(steps, sm_acc, col = prototype_label, fill = prototype_label)
  ) +
  facet_wrap(~prototype_label, ncol = 5) +
  geom_path(size = 2, show.legend = FALSE) +
  scale_color_manual(values = cividis(5)) +
  scale_fill_manual(values = cividis(5)) +
  labs(
    x = "Time step",
    y = "Average acceleration"
  )

vel_plot / acc_plot
```

```{r, include = FALSE}
export_ggsave("figure_11_velocity.pdf", width = 8, height = 3.5)
```



# Figure 12: Impact of trial design

## Retrieve and prepare data from OSF
```{r}
# Setup folder in working directory where design factors data should be stored
dir.create("design_factors_data")

# Download design factors data using OSF file links and read them into R
# (if files already exist, download will be skipped and data will still be loaded)
design_factors_raw <-
  c(
    "1" = "https://osf.io/7vrkz/",
    "2" = "https://osf.io/5hcju/",
    "3" = "https://osf.io/7bfhz/"
  ) %>%
  map_dfr(
    ~ osf_retrieve_file(.) %>% 
      osf_download(path = "design_factors_data", conflicts = "skip"),
    .id = "experiment"
  ) %>%
  pull(local_path, name = "experiment") %>%
  map_dfr(read_csv, .id = "experiment")

# Filter and label design factors data
design_factors_raw <-
  design_factors_raw %>%
  mutate(
    Typicality = factor(Condition, levels = c("Typical", "Atypical")),
    Manipulation = str_to_title(group),
    Manipulation = factor(
      Manipulation, 
      levels = c(
        "Click", "Touch", "Default", "Slow", "Static","Dynamic",
        "Initmax", "Rtmax"
        ),
      labels = c(
        "Click", "Touch", "Default", "Slow", "Static","Dynamic",
        "Timed: Initiation", "Timed: Response"
        )
    )
    
  ) %>%
  filter(correct == 1)
```

## Preprocess data
```{r}
# Import mouse-tracking data
mt_data <- mt_import_mousetrap(
  design_factors_raw,
  xpos_label = c("xpos_initial_phase", "xpos_get_response"),
  ypos_label = c("ypos_initial_phase", "ypos_get_response"),
  timestamps_label = c("timestamps_initial_phase", "timestamps_get_response")
)

# Preprocess mouse-tracking data
mt_data <- mt_data %>%
  mt_remap_symmetric() %>%
  mt_align_start() %>%
  mt_time_normalize()
```

## Create plot
```{r, fig.asp = 0.443}
# Create plots per experiment
p1 <-
  mt_data %>%
  mt_subset(experiment == 3) %>%
  mt_plot(
    use = "tn_trajectories",
    wrap_var = "Manipulation", wrap_ncol = 4,
    alpha = .025
  ) +
  coord_cartesian(xlim = c(-990, 990), ylim = c(-50, 970), expand = FALSE) +
  labs(x = NULL, y = NULL) + 
  theme(axis.text.x = element_blank())

p2 <-
  mt_data %>%
  mt_subset(experiment == 1) %>%
  mt_plot(
    use = "tn_trajectories",
    wrap_var = "Manipulation", wrap_ncol = 2,
    alpha = .025
  ) +
  coord_cartesian(xlim = c(-990, 990), ylim = c(-50, 970), expand = FALSE) +
  labs(x = NULL, y = NULL)

p3 <-
  mt_data %>%
  mt_subset(experiment == 2) %>%
  mt_plot(
    use = "tn_trajectories",
    wrap_var = "Manipulation", wrap_ncol = 2,
    alpha = .025
  ) +
  coord_cartesian(xlim = c(-990, 990), ylim = c(-50, 970), expand = FALSE) +
  labs(x = NULL, y = NULL) + 
  theme(axis.text.y = element_blank())

((p1) / (p2 | p3)) +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(size = 10))
```

```{r, include = FALSE}
export_ggsave(
  "figure_12_designfactors.png",
  width = 15.6, height = 6.9, dpi = 600, unit = "cm"
  )
```

