---
title: "KH2017 reanalysis following Dale et al. (2007)"
author: "Pascal J. Kieslich & Felix Henninger"
output:
  pdf_document: 
    fig_height: 4
    fig_width: 4
---

# References
Data from: Kieslich, P. J., & Henninger, F. (2017). Mousetrap: An integrated, open-source mouse-tracking package. _Behavior Research Methods, 49_(5), 1652-1667.
https://doi.org/10.3758/s13428-017-0900-z

Analyses following: Dale, R., Kehoe, C., & Spivey, M. J. (2007). Graded motor responses in the time course of categorizing atypical exemplars.
_Memory & Cognition, 35_(1), 15-28.
https://doi.org/10.3758/BF03195938



# General preparations

## Load libraries
```{r,message=FALSE}
library(mousetrap)
library(ggplot2)
library(afex)
```

## Custom ggplot2 theme
```{r}
theme_set(theme_classic()+ 
  theme(
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    axis.text = element_text(colour = "black"),
    panel.border = element_rect(colour = "black", fill=NA)
  ))
```



# Data preparation

## Read in merged raw data
```{r}
raw_data <- KH2017_raw
```

## Filter trials
```{r}
# Look at percent of correct responses per Condition
with(raw_data,table(Condition, correct)/c(table(Condition)))

# Test difference using chi-squared test
chisq.test(with(raw_data,table(Condition, correct)))

# Exlude incorrect trials
raw_data <- subset(raw_data, correct==1)

# Check that each subject has at least one typical and atypical trial left
table(with(raw_data,table(subject_nr,Condition))>0)
```

## Import mouse-tracking data
```{r}
mt_data <- mt_import_mousetrap(raw_data)
```

## Trajectory preprocessing
```{r}
mt_data <- mt_remap_symmetric(mt_data)
mt_data <- mt_align_start(mt_data)
mt_data <- mt_time_normalize(mt_data)
```



# Analyses

## Analysis of time-normalized trajectories

### Plot average time-normalized tajectories
```{r}
mt_plot_aggregate(mt_data, use = "tn_trajectories", points = TRUE,
  x = "xpos", y = "ypos", color = "Condition", subject_id = "subject_nr")+
  scale_color_brewer(type="qual",palette = "Set1" )+
  theme(legend.position=c(.2,.2))
```


### Compare time normalized x-positions for each step
```{r}
# Aggregate time-normalized trajectories per condition
# separately per subject
av_tn_trajectories <- mt_aggregate_per_subject(mt_data,
  use="tn_trajectories",use2_variables="Condition",subject_id="subject_nr")

# Plot aggregate trajectories with standard errors
# (note that mean_se does not take into account within subjects design)
ggplot(av_tn_trajectories,aes(x=steps,y=xpos,group=Condition))+
  stat_summary(geom = "ribbon",fun.data=mean_se,alpha=.2)+
  geom_line(aes(color=Condition),stat="summary",fun.y="mean")+
  scale_color_brewer(type="qual",palette = "Set1" )+
  theme(legend.position=c(.2,.2))


### Paired t-tests on coordinates
xpos_t_tests <- 
  with(av_tn_trajectories,
       sapply(unique(steps),function(i){
         t.test(xpos[Condition=="Atypical" & steps==i],
                xpos[Condition=="Typical" & steps==i],
                paired = TRUE)$p.value})
       )

# Retrieve all significant t-tests
which(xpos_t_tests<.05)

# Number of significant t-tests
sum(xpos_t_tests<.05,na.rm=TRUE)

# Number of adjacent significant t-tests (minus 1)
table(diff(which(xpos_t_tests<.05)))
```


### Analysis of binned time-normalized trajectories
```{r,message=FALSE,warning=FALSE}
mt_data <- mt_average(mt_data,use="tn_trajectories",save_as="av_tn_trajectories",
  av_dimension="steps",intervals = c(0.5,33.5,67.5,101.5))

mt_aggregate(mt_data,use = "av_tn_trajectories",
  use2_variables = "Condition",subject_id = "subject_nr")

mt_plot_aggregate(mt_data,use = "av_tn_trajectories", points=TRUE,
  color = "Condition",subject_id = "subject_nr")+
  scale_color_brewer(type="qual",palette = "Set1" )


av_tn_trajectory_bins <-  mt_aggregate_per_subject(mt_data,use="av_tn_trajectories",
  use2_variables="Condition",subject_id="subject_nr")

av_tn_trajectory_bins$bin <- factor(av_tn_trajectory_bins$steps)

aov_ez(data=av_tn_trajectory_bins,id="subject_nr",dv="xpos",within = c("Condition","bin"),
       anova_table = list(es=c("ges","pes"),correction=c("none")))

aov_ez(data=av_tn_trajectory_bins,id="subject_nr",dv="xpos",within = c("Condition","bin"),
       anova_table = list(es=c("ges","pes"),correction=c("GG")))
```


## Analysis of space normalized trajectories per subject

```{r,message=FALSE,warning=FALSE}
mt_data <- mt_align_start_end(mt_data,save_as="sn_trajectories",
  start = c(0,0), end = c(-1,1))
mt_data <- mt_average(mt_data,use="sn_trajectories",
  interval_size = 500,max_interval=1500)

mt_aggregate(mt_data,use="av_trajectories",
  use2_variables="Condition",subject_id="subject_nr")

mt_plot_aggregate(mt_data,use = "av_trajectories", points=TRUE,
  color = "Condition",subject_id = "subject_nr")+
  scale_color_brewer(type="qual",palette = "Set1" )

av_sn_trajectory_bins <-  mt_aggregate_per_subject(mt_data,use="av_trajectories",
  use2_variables="Condition",subject_id="subject_nr")

# Check that each subject has observations in each category
table(table(av_sn_trajectory_bins$subject_nr))

av_sn_trajectory_bins$bin <- factor(av_sn_trajectory_bins$timestamps)

aov_ez(data=av_sn_trajectory_bins,id="subject_nr",dv="xpos",within = c("Condition","bin"),
       anova_table = list(es=c("ges","pes"),correction=c("none")))

aov_ez(data=av_sn_trajectory_bins,id="subject_nr",dv="xpos",within = c("Condition","bin"),
       anova_table = list(es=c("ges","pes"),correction=c("GG")))
```


## Distributional analyses
```{r}
# Compute trial indices
mt_data <- mt_measures(mt_data)

# Standardize measures per participant
mt_data <- mt_standardize(mt_data,
  use_variables = c("MAD","AUC","AD"),within = "subject_nr",prefix = "z_")

# Merge data
results <- merge(mt_data$data,mt_data$measures,by="mt_id")

# Distribution
qplot(x=z_MAD,data=results,bins=50)+facet_grid(Condition~.)
qplot(x=z_AUC,data=results,bins=50)+facet_grid(Condition~.)
qplot(x=z_AD,data=results,bins=50)+facet_grid(Condition~.)

# Calculate bimodality coefficient
mt_check_bimodality(mt_data,use_variables = c("z_MAD","z_AUC","z_AD"),
  grouping_variables = "Condition", methods = c("BC"))

```


## Sample entropy
```{r}
# Calculate sample entropy (for lag 3)
mt_data <- mt_sample_entropy(mt_data,use="tn_trajectories",m=3,dimension="xpos")

# Aggregate per participant
agg_sample_entropy <- mt_aggregate_per_subject(mt_data, subject_id = "subject_nr",
  use_variables = "sample_entropy", use2_variables = "Condition")

# Test the difference
mt_aggregate(mt_data, subject_id = "subject_nr",
  use_variables = "sample_entropy", use2_variables = "Condition")
t.test(sample_entropy~Condition, data=agg_sample_entropy, paired=TRUE)
```



# Plot for left- and rightward responses

## Data preparation
```{r}
mt_data <- mt_import_mousetrap(raw_data)
mt_data <- mt_remap_symmetric(mt_data,remap_xpos = "no")
mt_data <- mt_align_start(mt_data)
mt_data <- mt_time_normalize(mt_data)

# code response side
mt_data$data$side <- ifelse(mt_data$data$response==mt_data$data$CategoryLeft,"left","right")
```


## Plot average time-normalized tajectories
```{r,fig.width=7}
mt_plot_aggregate(mt_data, use = "tn_trajectories", points = TRUE,
  x = "xpos", y = "ypos", color = "Condition",
  linetype = "side")+
  scale_color_brewer(type="qual",palette = "Set1" )+
  scale_linetype_manual(values=c("solid","solid"),guide=FALSE)+
  theme(legend.position=c(.15,.2))
```

